---
title: "possible_cars"
author: "Danielle Sclafani"
date: "2/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
library(ggpubr)
```

## Recreating Possible Cars from ESM 262 Examples


#What varies- 4 car types
Imagine with have 3 different car types - and we know how often each occurs:

* car A  mass 20000 kg, area 22 m^2^

* car B mass 65000 kg, area 30 m^2^

* car C mass 38000 kg area 22 m^2^

* car D mass 5000 kg area 10m^2^

mean highway speed is 100 km/hr (must convert to miles per hour)

```{r}
#source 
#aut_power_gen = function(mass, speed, A, crolling=0.015, g=9.8, p_air= 1.2,cdrag = 0.3){
#  result = crolling*mass*g*speed + 1/2*A*p_air*cdrag*speed^3
 # return(result)

#creating the data frame of the cars, their masses, and surface areas
possible_cars = data.frame(name = c("A", "B", "C", "D"), mass = c(20000, 65000, 38000, 5000), area = c(22 , 30, 22, 10))

speed_base = 100*0.28

#input this one speed into the formula to get the power- using the code below it will add the power as another column to the data frame 
possible_cars$power = aut_power_gen(mass = possible_cars$mass, speed = speed_base, A= possible_cars$area)

#showing the results
#ggplot(possible_cars, aes(x=mass, y=power, fill=as.factor(area)))+geom_col()+labs(y="Power W", x="Mass (kg)", fill="Area (m2)")


```

# Building a highway

What could be the total power consumed if there are 100 cars on this highway each hour,
they are travelling at a range of speeds - mean is 100km/hr and speeds tend to
be log-normally distributed)

How would the total power consumed vary by car
So if all cars are car A; OR all cars are car B OR all cars are car C

We will use *sample* here to generate speeds for our 100 cars and look at different ways to
repeat power calculation for different cars

* repeating by hand
* *pmap* for repetition - a efficient way of looping in R
* *for* for repetition - a more standard way of looping available in many langugaes

# First lets do it 'by hand'
```{r}
# assume a log normal distribution of speeds with mean 100km/hr, and standard deviation that is 10% of the mean

# recall our function needs speed in m/s not km/hr so we will also do a conversion
# lets get a sample of a 200 speeds- we could also do this by actually measuring speeds

nsample = 200
mean_speed = log(100*0.277)

speeds = rlnorm(mean=mean_speed, sd=0.1*mean_speed, nsample)
summary(speeds)

plot(density(speeds), ylab="Distribution of Speeds in (m/s)")
```

```{r}
#doing the calc by hand - finding the power for each car one at a time
# we could do it by hand - input the various speeds into each formula to get one possibility of the power for that one speed

powerA = aut_power_gen(mass = possible_cars$mass[1], speed = speeds, A= possible_cars$area[1])
# lets look at what we get
summary(powerA)

# next car (row 2)
powerB = aut_power_gen(mass = possible_cars$mass[2], speed = speeds, A= possible_cars$area[2])
# lets look at what we get
summary(powerB)

# next car (row 3)
powerC = aut_power_gen(mass = possible_cars$mass[3], speed = speeds, A= possible_cars$area[3])
# lets look at what we get
summary(powerC)

#the new car (row 4)
powerD = aut_power_gen(mass = possible_cars$mass[4], speed = speeds, A= possible_cars$area[4])

#want to bind these 4 data points into one dataframe

powerall_cars= cbind.data.frame(powerA, powerB, powerC, powerD) #this gives you the different powers for each car at varying speeds. format of these data frame not that helpful so want to make the columns into rows

powerall_cars1 = powerall_cars %>% pivot_longer(cols = everything(), names_to="car", values_to="power")

#method1_plot = ggplot(powerall_cars1, aes(car,power, fill=car))+geom_boxplot()+ggtitle("By Hand")
#method1_plot
```

#Second method is using the built in tools in R

```{r}
#speed distribution is the same, so are the cars

#use pmap to run each car for all speeds- sample = 100 speeds along the log distribution

powerall_cars2 = pmap(list(A=possible_cars$area, mass= possible_cars$mass),aut_power_gen, speed= speeds)

#need to make the pmap function into a dataframe and add column names- the same as possible_cars$name

powerall_cars2 = as.data.frame(powerall_cars2, col.names = possible_cars$name)

# apply family of functions does this to - FYI
# what mapply does is run the function for each row in parameters listed, using values for other parameters listed in MoreArgs EACH time - a column for row in parameter list is returned

powerall_cars2b = mapply(FUN = aut_power_gen, A = possible_cars$area, mass = possible_cars$mass, MoreArgs = list(speed=speeds))

#add column names
colnames(powerall_cars2b) = possible_cars$name

head(powerall_cars2)
head(powerall_cars2b)

#turn column into rows for easier plotting
powerall_cars2_r = powerall_cars2 %>% pivot_longer(cols = everything(), names_to="car", values_to="power")
head(powerall_cars2_r)

#lets look at this
method2_plot = ggplot(powerall_cars2_r, aes(car,power, fill=car))+geom_boxplot()+ggtitle("pmap")
method2_plot
```


```{r}
#looking at the mean of 
powerall_cars %>% map(mean)
powerall_cars2 %>% map(mean)
```


# using the classic "for" looping function

```{r}
# using the same as above
possible_cars
speeds

# how do we run each car for all speeds 
# we use a for loop to cycle through
# we need to create a data frame to store results - as above
# one column for each car and one row for each speed

powerall_cars3 = as.data.frame(matrix(nrow = length(speeds), ncol = nrow(possible_cars)))
#values should be NA because do not initialize it

#how many cars are there
nrow(possible_cars) # when return this it says 4 cars, which is correct
# for loops use an index - in this case "i" but you could use anything - it repeats
# anything between the {} for each values of i between 1 and nrow(possible_car) (which is 4 in our case)

# index in to a matrix (like powerall3) is by row and column powerall3[2,5] is 2nd row and 5th column
for (i in 1:nrow(possible_cars)){
  powerall_cars3[,i] = aut_power_gen(A=possible_cars$area[i], m=possible_cars$mass[i], speed=speeds)
}

head(powerall_cars3) #returns column names as V1, V2, V3, V4
#add column names
colnames(powerall_cars3)=possible_cars$name # this names them car A, B, C, D

#turn columns to rows
powerall_cars3r = powerall_cars3 %>%pivot_longer(cols = everything(), names_to="car", values_to = "power")
head(powerall_cars3r)

#look at it
method3_plot = ggplot(powerall_cars3r, aes(car,power, fill=car))+geom_boxplot()+ggtitle("For Loop")
method3_plot 

ggarrange(method1_plot, method2_plot, method3_plot, nrow=3)

```




