---
title: "automobile_power_gen"
author: "Danielle Sclafani"
date: "2/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Autombobile Power Generation Funciton
Contract
*Input*: Vehicle speed, V (m/s), vehicle mass, m  (kg), surface area of the car, A (m2)
*Output*: Power generation Pb (W)
*Parameters*: acceleration due to gravity, g (9.8 m/s2), density of air, p_air (1.2 kg/m3), crolling (0.015), cdrag (0.3)

The function will generate the power of the automobile needed to maintain the vehicles speed

$Pb = crolling * m * g * V + 1/2 * A * p_air * cdrag * V^3$

### Write a function to compute power, given a truck of m=31752 kg (parameters for a heavy truck) for a range of different highway speeds (30 m/s, 80 km/hr, 120 km/hr, 40 km/hr, 5km/hr ) The surface area is 16 m2

```{r}
#aut_power_gen = function(mass, speed, A, crolling=0.015, g=9.8, p_air= 1.2,cdrag = 0.3){
 # result = crolling*m*g*speed + 1/2*A*p_air*cdrag*speed^3
 # return(result)}

aut_power_gen(31752, 30, 16)
```


```{r}
#converting m/s to km/hr
#80 km/hr, 120 km/hr, 40 km/hr, 5km/hr

#80 km/hr
80/3.6 
#= 22.22 m/s
aut_power_gen(31752, 22.22, 16)
#135308.3

#120 km/hr
120/3.6 
#= 33.33 m/s
aut_power_gen(31752, 33.33, 16)
#262203.9

#40 km/hr
40/3.6 
#= 11.11 m/s
aut_power_gen(31752, 11.11, 16)
#55805.85

#5 km/hr
5/3.6 
#= 1.39 m/s
aut_power_gen(31752, 1/39, 16)
#119.6807

```

#making a data frame
```{r}

power = aut_power_gen(m=31752, speed=c(22.22, 33.33, 11.11, 1.39), A=16)

power_speed = data.frame(power, 
                         speed = c (22.22, 33.33, 11.11, 1.39))

ggplot(data = power_speed, aes(x = speed, y = power)) + geom_line()
```

## Using a lighter car
```{r}

power_light = aut_power_gen(m=21752, speed=c(22.22, 33.33, 11.11, 1.39), A=16)

power_speed_light = data.frame(power_light, 
                         speed = c (22.22, 33.33, 11.11, 1.39))

ggplot() + geom_line(data = power_speed_light, aes(x=speed, y = power_light)) + geom_line(data = power_speed, aes(x = speed, y = power))
```

